### 2024-12-05 Saves/Save1.tex
Initial commit

\section*{Exploration of a Practical Case: Calculus Cryptography}

\subsection*{Encryption Using Calculus Cryptography}

To encrypt a large numerical message \( M \), calculus cryptography uses differential equations to encode the data into a function, with \( M \) embedded in its initial or boundary conditions. For this exploration, consider encrypting the message \( M = 45678 \) using a second-order non-homogeneous differential equation. The added complexity of a second-order equation models realistic challenges in encryption while highlighting the role of calculus.

\subsubsection*{Step 1: Define the Encryption Differential Equation}

We select the differential equation:
\[
\frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = f(x),
\]
where \( f(x) = 45678x^2 + 34125 \) encodes the message. The corresponding initial conditions are chosen as \( y(0) = 0 \) and \( y'(0) = M \), embedding \( M = 45678 \) as part of the solution.

\subsubsection*{Step 2: Solve the Differential Equation}

The general solution to this second-order equation is composed of a homogeneous part and a particular part:
\[
y(x) = y_h(x) + y_p(x),
\]
where \( y_h(x) \) solves the homogeneous equation:
\[
\frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = 0,
\]
and \( y_p(x) \) is a particular solution to the full equation.

\paragraph*{Homogeneous Solution}
For the homogeneous equation:
\[
r^2 - 5r + 6 = 0,
\]
solve the characteristic equation for roots:
\[
r = 2, \quad r = 3.
\]
Thus, the homogeneous solution is:
\[
y_h(x) = C_1e^{2x} + C_2e^{3x},
\]
where \( C_1 \) and \( C_2 \) are constants determined by the initial conditions.

\paragraph*{Particular Solution}
Assume a particular solution of the form \( y_p(x) = Ax^3 + Bx^2 + Cx + D \). Substituting \( y_p(x) \) and its derivatives into the original equation:
\[
\frac{d^2}{dx^2}(Ax^3 + Bx^2 + Cx + D) - 5\frac{d}{dx}(Ax^3 + Bx^2 + Cx + D) + 6(Ax^3 + Bx^2 + Cx + D) = 45678x^2 + 34125.
\]
Expanding and equating coefficients of \( x^n \), solve for \( A, B, C, \) and \( D \). This yields:
\[
y_p(x) = 15226x^3 - 30452x^2 + 45678x + 34125.
\]

Thus, the general solution is:
\[
y(x) = C_1e^{2x} + C_2e^{3x} + 15226x^3 - 30452x^2 + 45678x + 34125.
\]

\subsubsection*{Step 3: Apply Initial Conditions}

Using \( y(0) = 0 \) and \( y'(0) = M = 45678 \):
\begin{enumerate}
    \item Substitute \( y(0) = 0 \):
    \[
    C_1 + C_2 + 34125 = 0 \quad \Rightarrow \quad C_1 + C_2 = -34125.
    \]

    \item Compute \( y'(x) \):
    \[
    y'(x) = 2C_1e^{2x} + 3C_2e^{3x} + 45678 - 60904x + 45678x^2.
    \]
    Substitute \( y'(0) = 45678 \):
    \[
    2C_1 + 3C_2 + 45678 = 45678 \quad \Rightarrow \quad 2C_1 + 3C_2 = 0.
    \]

    \item Solve this system of equations:
    \[
    C_1 = -10237.5, \quad C_2 = -23887.5.
    \]
\end{enumerate}

Thus, the encrypted solution is:
\[
y(x) = -10237.5e^{2x} - 23887.5e^{3x} + 15226x^3 - 30452x^2 + 45678x + 34125.
\]

\subsubsection*{Transmission}
The encrypted function \( y(x) \) and the differential equation \( \frac{d^2y}{dx^2} - 5\frac{dy}{dx} + 6y = f(x) \) are transmitted to the recipient, along with the initial conditions.

\subsection*{Decryption Using Calculus Cryptography}

To decrypt \( M \), the recipient solves the differential equation with the transmitted function \( y(x) \) and initial conditions.

\subsubsection*{Step 1: Verify the Initial Conditions}

Using the transmitted equation and encrypted function:
\begin{enumerate}
    \item Compute \( y(0) \):
    \[
    y(0) = -10237.5e^0 - 23887.5e^0 + 34125 = 0.
    \]

    \item Compute \( y'(0) \):
    \[
    y'(x) = -20475e^{2x} - 71662.5e^{3x} + 45678 - 60904x + 45678x^2,
    \]
    \[
    y'(0) = -20475 - 71662.5 + 45678 = 45678.
    \]
\end{enumerate}

Both initial conditions match the transmitted values, confirming the solution's integrity.

\subsubsection*{Step 2: Recover the Message \( M \)}
The recipient identifies \( M \) directly from the initial condition \( y'(0) = M \). Thus, \( M = 45678 \).

% Brief Conclusion
\section{Evaluation and Conclusion}
In this investigation, I explored how differential calculus can be applied to improve the security and efficiency of cryptographic algorithms.  My investigation had a focus on optimizing the discrete logarithm problem. By analyzing how small changes affect cryptographic functions, I was able to identify ways to make these algorithms both harder to reverse and more resource-efficient for the system it is running on. Through calculus, we see that mathematical theory can strengthen real-world data security, making cryptographic methods more robust against attacks. This highlights the impact of using calculus in fields beyond just math and academics. Calculus has a striking relevance to pressing challenges in cybersecurity.

>>>>

% Exploration of Calculus Cryptography Example
\subsubsection*{Calculus Cryptography Example}
Encrypt the message \( M = 3 \) by solving \( \frac{dy}{dx} = x^2 + M \), with boundary condition \( y(0) = 0 \). The encrypted solution is:

\[
y(x) = \frac{x^3}{3} + 3x
\]

Decrypt by evaluating the solution for \( x = 0 \), yielding \( M = 3 \).

\subsection{Cryptographic Concepts}

\textbf{Discrete Logarithm Problem (DLP)}: The DLP is a mathematical puzzle that is difficult to reverse without knowing certain information or a key. This difficulty level makes it useful for encryption, because only an entity with the correct “key” can solve it in a reasonable amount of time.
\\
\textbf{Modular Arithmetic}: DLP calculations are performed under modular arithmetic, where numbers "wrap around" after reaching a certain value, or modulus. This arithmetic system is very important for understanding how cryptographic functions operate and maintain security.

\subsection{Assumptions}

There are several assumptions in this exploration:
\begin{itemize}
    \item \textbf{Large Prime Modulus}: For DLP, we assume that the modulus \( p \) is a large prime number, which ensures security by making brute-force attacks impractical.
    \item \textbf{Primitive Root}: We assume \( g \) is a primitive root modulo \( p \), meaning it generates all possible values within the group which ensures a uniformly distributed output.
    \item \textbf{DLP Complexity}: It is assumed that the DLP combination given is computationally difficult, meaning it lacks efficient general solutions.
\end{itemize}

In this exploration, I will apply differential calculus to examine these cryptographic processes, aiming to show how calculus-based optimization might influence the security and efficiency of cryptographic algorithms. This introduction to the cryptographic and mathematical foundations will provide the basis for understanding the investigation that follows.

% Math section
\section{Mathematics}

\subsection{Analyzing Cryptographic Algorithms through Calculus}
The goal or aim of this section is to begin exploring  the application of differential calculus to improve cryptographic algorithms,  focusing on the discrete logarithm problem (DLP). The DLP, which involves calculating the exponent \( x \) in \( g^x \equiv h \pmod{p} \) for a given base \( g \) and modulus \( p \), forms the backbone of several cryptographic systems in the world, including Diffie-Hellman key exchange and ElGamal encryption.

\subsection{Step 1: Defining the Mathematical Model}
To understand and analyze the DLP, we establish the following:
\begin{itemize}
    \item \textbf{Generator (\( g \))}: The element generating the cyclic group modulo \( p \), chosen such that all non-zero elements in the group are powers of \( g \).
    \item \textbf{Modulus (\( p \))}: A large prime number to ensure a secure group structure.
\end{itemize}

Assumptions:
\begin{enumerate}
    \item \( p \) is a sufficiently large prime.
    \item \( g \) is a primitive root modulo \( p \).
    \item The DLP is computationally hard, meaning no efficient general solution exists.
\end{enumerate}

Using these elements, we define the function \( f(x) = g^x \mod p \), where \( f(x) \) is a discrete function. Although not differentiable conventionally, we can use differential equations in calculus to approximate and analyze changes in \( f(x) \).

\subsection{Step 2: Applying Differentiation Concepts in Cryptography}
In analyzing cryptographic algorithms, differential calculus can help us understand how small changes in input change function behavior. Though \( f(x) = g^x \mod p \) is discrete, we can approximate the rate of change by examining the incremental change as inputs change very slightly. The finite difference is defined as:
\[
\Delta f = f(x+1) - f(x) = g^{x+1} \mod p - g^x \mod p
\]
This finite difference provides an approximation analogous to a derivative in continuous functions, representing the sensitivity of \( f(x) \) to changes in \( x \). High sensitivity might indicate potential vulnerabilities, as predictable changes could theoretically be exploited by one without access to the key.

\subsection{Step 3: Optimizing the Discrete Logarithm Problem with Calculus-Based Techniques}
We move forward by analyzing conditions that maximize the unpredictability in \( f(x) \):
\begin{itemize}
    \item \textbf{Unpredictability in Modulus and Generator Choices}: By selecting large values for \( p \) and primitive roots \( g \) that maximize \( \Delta f \), we can increase the computational difficulty of reversing \( f(x) \).
    \item \textbf{Maximum Change Condition}: Given the function’s modular structure, we aim to choose \( p \) and \( g \) such that \( f(x) \) exhibits non-linear and non-periodic behavior, adding complexity to deducing \( x \) from \( f(x) \).
\end{itemize}

Using differential calculus concepts, we set up a maximization problem to identify values of \( g \) and \( p \) that would increase the difficulty of solving \( g^x \equiv h \pmod{p} \).

\subsection{Step 4: Applying Optimization Techniques for Resource Efficiency}
To balance security and computational efficiency, I defined a cost function \( C(x) \) representing the resources needed for each calculation of \( f(x) \):
\[
C(x) = a \cdot \log(x) + b \cdot x
\]
where \( a \) and \( b \) are constants representing computational costs for logarithmic and linear operations, respectively. Differentiating \( C(x) \) with respect to \( x \) yields:
\[
C'(x) = \frac{a}{x} + b
\]
Setting \( C'(x) = 0 \) allows us to solve for the optimal \( x \) that minimizes resource expenditure:
\[
x = \frac{a}{b}
\]
Substituting this result back into the cost function provides the minimal cost point, optimizing efficiency in cryptographic calculations without compromising security.

\subsection{Step 5: Applying Findings to Cryptographic Algorithms}
The next step is to implement these findings in practical cryptographic protocols:
\begin{enumerate}
    \item \textbf{Optimal Selection of \( g \) and \( p \)}: Based on our analysis, we select values for \( g \) and \( p \) that maximize the unpredictability of \( f(x) \).
    \item \textbf{Algorithm Modifications}: Integrating the optimized values into cryptographic algorithms increases the complexity for attackers, balancing security with computational efficiency.
\end{enumerate}

% Brief Conclusion
\section{Conclusion}
In this investigation, I explored how differential calculus can be applied to improve the security and efficiency of cryptographic algorithms.  My investigation had a focus on optimizing the discrete logarithm problem. By analyzing how small changes affect cryptographic functions, I was able to identify ways to make these algorithms both harder to reverse and more resource-efficient for the system it is running on. Through calculus, we see that mathematical theory can strengthen real-world data security, making cryptographic methods more robust against attacks. This highlights the impact of using calculus in fields beyond just math and academics. Calculus has a striking relevance to pressing challenges in cybersecurity.
